---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Evolution of Legends: A Comparative Analysis of Coldplay and Metallica's Musical Progression in the Context of Broader Industry Trends"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Joshua-Connor Knapp^[__Contributions:__  \\newline _I'm not a snob, objectively, my tatse in music is just better than yours._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Flex Offenders, Local Garage Gym, South Africa" # First Author's Affiliation
Email1: "joshuaconnok\\@gmail.com" # First Author's Email address

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#ELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This report presents a comparative analysis of the musical progression of two iconic bands, Coldplay and Metallica, over their multi-decade careers. Utilizing data from Spotify and supplementary sources such as the Billboard Top 100 charts, the study examines key musical attributes including tempo, energy, valence, danceability, and popularity. The analysis explores and contextualizes the bands' stylistic changes within broader music industry trends, identifying correlations and contrasts with mainstream music. This comprehensive examination provides insights into how Coldplay and Metallica have evolved over time, their unique contributions to the music landscape, and their alignment with or divergence from prevailing musical trends.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

library(tidyverse)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

library(readr)
library(tidyverse)

coldplay <- read_csv("data/Coldplay.csv")
metallica <- read_csv("data/metallica.csv")
spotify <- read_rds("data/Broader_Spotify_Info.rds")
billboard_100 <- read_rds("data/charts.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

Coldplay and Metallica boast careers spanning over decades, during which they have made significant contributions to the music industry with their unique styles and consistent evolution. This analysis aims to explore various aspects of their musical journeys by examining their discographies and analyzing how their musical attributes have evolved over time. By leveraging comprehensive data from Spotify, this report will delve into key metrics such as tempo, energy, valence, and danceability. Additionally, the analysis will contextualize these bands' stylistic changes within broader music industry trends, using supplementary data sources such as the Billboard Top 100 charts.To provide a detailed comparison and identify patterns, we will ensure that the analysis only considers studio recordings, filtering out live performances and demo versions. This approach will maintain the focus on like-for-like comparisons, allowing us to draw meaningful insights into each band's musical evolution.

# Data  {-}

The dataset provided for the analysis of Coldplay and Metallica's musical progression includes a range of attributes that capture various aspects of their songs. Each track is characterized by its duration, measured in seconds, and the release_date, which indicates the date the album was released. The album_name provides the name of the album to which the track belongs. The explicit attribute indicates whether the song contains explicit lyrics, which can impact its audience and distribution. The popularity score, ranging from 0 to 100, measures the song's popularity on Spotify.Several attributes describe the musical qualities of the tracks. Acousticness measures the confidence that a track is acoustic, with values ranging from 0.0 to 1.0. Danceability assesses how suitable a track is for dancing, based on elements like tempo and beat strength. Energy reflects the intensity and activity level of the song, a higher energy level indicating more dynamic and engaging tracks. Instrumentalness predicts the likelihood of a track containing no vocal content, with higher values indicating more instrumental tracks. Liveness detects the presence of an audience in the recording, with values above 0.8 suggesting a live performance. The dataset also includes loudness, which measures the perceived loudness of the track in LUFS, providing a more accurate reflection of how humans perceive loudness. Speechiness detects the presence of spoken words, with higher values indicating more speech-like content. Tempo measures the speed of the track in beats per minute (BPM), and valence describes the musical positiveness conveyed by the track, with higher values indicating more positive, happy-sounding tracks. Together, these attributes offer a comprehensive view of the musical characteristics and evolution of Coldplay and Metallica over their long careers.

# Comparison

<!-- The following is a code chunk. It must have its own unique name (after the r), or no name. After the comma follows commands for R which are self-explanatory. By default, the code and messages will not be printed in your pdf, just the output: -->

As can be seen from the plots below, while Metallica has reached greater a level of popularity for one of their albums, the self titled "Metallica", this popularity is inconsistent with other albums falling far below this level. Coldplay see's much more consistent levels of popularity across its album releases. Metallica Has also released more albums than Coldplay.

```{r Figure1, echo=FALSE, fig.align='center', fig.cap="Coldplay Album Popularity \\label{Figure1}", fig.ext='png', fig.height=4, fig.width=6, warning=FALSE}

# Using functions to filter out all non-studio songs...

coldplay_studio <- studio_filter(coldplay)

# ... and plotting as box and whiskers

c_plot <- box_plotter(coldplay_studio, "popularity", title = "Popularity by album (Coldplay)")
c_plot

```


```{r Figure2, echo=FALSE, fig.align='center', fig.cap="Metallica Album Popularity \\label{Figure2}", fig.ext='png', fig.height=4, fig.width=6, warning=FALSE}

# Using functions to filter out all non-studio songs...

metallica_studio <- studio_filter(metallica)

# ... and plotting as box and whiskers

m_plot <- box_plotter(metallica_studio, "popularity", title = "Popularity by album (Metallica)")
m_plot
```

To contextualize the musical evolution of Coldplay and Metallica within broader music industry trends, I utilized supplementary data sources, including the Billboard Top 100 charts spanning several decades. This analysis involved comparing the attributes of Coldplay and Metallica's tracks—such as tempo, energy, valence, danceability, and popularity—with trends observed in mainstream music over the years. By examining changes in these musical attributes and their alignment with or divergence from the broader trends in the Billboard Top 100, I identified correlations and contrasts in stylistic changes. This comparative approach allowed for a deeper understanding of how Coldplay and Metallica's musical progression aligns with or stands apart from the prevailing trends in popular music, shedding light on their unique contributions to the industry and their ability to adapt to or influence broader musical movements.

```{r Figure3, echo=FALSE, fig.align='center', fig.cap="Evolution of Music over time \\label{Figure3}", fig.height=5, fig.width=6, warning=FALSE, fig.ext='png'}

# First wrangling the Spotify and billboard data a little bit and joining them...

billboard_100 <- billboard_100 %>% 
    rename(name = song)

mainstream_music <- inner_join(spotify, billboard_100, by = c("name", "artist"))
mainstream_music <- mainstream_music %>% 
    rename(release_date = date)

# .. then feeding it to my attributes plotting function

attributes <- band_attributes(metallica_studio, coldplay_studio, mainstream_music, band_name1 = "Metallica", band_name2 = "Coldplay", band_name3 = "Billboard100")
print(attributes)
```

Plotting heatmaps...

```{r Figure4, echo=FALSE, fig.align='center', fig.cap="Heatmap of Metallicas musical attributes\\label{Figure4}", fig.height=3, fig.width=5, warning=FALSE, fig.ext='png'}

m_heatmap <- heat_mapper(metallica_studio, title = "Heatmap for Metallica Songs")
m_heatmap

```

```{r Figure5, echo=FALSE, fig.align='center', fig.cap="Heatmap of Coldplays musical attributes\\label{Figure5}", fig.height=3, fig.width=5, warning=FALSE, fig.ext='png'}

c_heatmap <- heat_mapper(coldplay_studio, title = "Heatmap for Coldplay Songs")
c_heatmap

```


 
# Conclusion




Graph Analysis
1. Popularity by Album
The two boxplots represent the popularity of each album for both Metallica and Coldplay.

Metallica:

The popularity scores for Metallica's albums vary widely.
Their earlier albums such as "Master of Puppets," "Ride the Lightning," and "Kill 'Em All" show higher popularity scores with less variability.
Some newer albums like "Hardwired… to Self-Destruct" show lower popularity scores.
This suggests a strong start with a decline in popularity for later albums, though some recent works still maintain considerable popularity.
Coldplay:

Coldplay shows a different trend with a more consistent popularity score across albums.
Albums like "A Rush of Blood to the Head" and "Viva La Vida or Death and All His Friends" have higher popularity scores.
There is less variability in the popularity scores for Coldplay’s albums, suggesting a steadier fanbase.
2. Musical Progression
The line graphs depict the progression of various musical attributes (energy, valence, danceability, and duration) over time for Metallica and Coldplay, compared to the Billboard 100 trends.

Energy:

Metallica shows high energy levels throughout their career, with fluctuations but generally maintaining a high energy profile.
Coldplay shows a rising trend in energy over the years, indicating a shift towards more energetic music in their later works.
Valence:

Valence (musical positivity) for both bands shows interesting trends. Metallica generally has a lower valence compared to Coldplay.
Coldplay's valence has varied but shows a slight upward trend, indicating more positive or emotionally uplifting music in recent years.
Danceability:

Metallica maintains a lower danceability score, consistent with their genre and style.
Coldplay's danceability shows an increasing trend, aligning with their move towards more pop and rhythm-oriented music in recent albums.
Duration:

Both bands exhibit changes in song duration over the years. Metallica’s song durations have seen more fluctuation.
Coldplay's songs have become longer over time, which might correlate with more complex song structures or varied musical explorations.
3. Heatmap of Musical Attributes
The heatmaps show the correlation between different musical attributes for Metallica and Coldplay songs.

Metallica:

Positive correlations: Loudness and energy are highly correlated, as are acousticness and liveness.
Negative correlations: Danceability and acousticness, danceability and loudness, and instrumentalness and valence have notable negative correlations.
This suggests that Metallica's loud and energetic music tends to be less danceable and less acoustic.
Coldplay:

Positive correlations: Energy and loudness, acousticness and valence.
Negative correlations: Danceability and acousticness, liveness and acousticness.
This indicates that Coldplay's more energetic and loud songs tend to be less acoustic, and there is a distinct separation between acoustic and danceable tracks.
Report Summary
Longevity and Popularity:

Metallica’s popularity has seen significant highs and lows, with early albums being more popular. Coldplay has maintained a more consistent popularity across their discography.
Musical Evolution:

Metallica has maintained a high-energy profile, true to their heavy metal roots, but with decreasing valence and danceability in line with their genre.
Coldplay has evolved from a softer, alternative rock sound to incorporating more energetic and danceable elements in recent years, reflecting a trend towards mainstream pop.
Comparative Trends:

Compared to the Billboard 100, Metallica's music stands out with higher energy and lower valence, emphasizing their heavy metal identity.
Coldplay’s musical progression aligns more closely with mainstream trends, showing increasing energy and danceability, which might explain their consistent popularity.
Conclusion:

Metallica and Coldplay have both demonstrated impressive longevity but have taken very different musical journeys. Metallica's steadfast adherence to their heavy metal roots contrasts with Coldplay's adaptive style, which has embraced more modern and mainstream trends. This comparative analysis highlights the diverse ways in which bands can evolve and sustain their relevance in the music industry over decades.








