---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Evolution of Legends: A Comparative Analysis of Coldplay and Metallica's Musical Progression in the Context of Broader Industry Trends"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Joshua-Connor Knapp^[__Contributions:__  \\newline _I'm not a snob, objectively, my tatse in music is just better than yours._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Flex Offenders, Local Garage Gym, South Africa" # First Author's Affiliation
Email1: "joshuaconnok\\@gmail.com" # First Author's Email address

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#ELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This report presents a comparative analysis of the musical progression of two iconic bands, Coldplay and Metallica, over their multi-decade careers. Utilizing data from Spotify and supplementary sources such as the Billboard Top 100 charts, the study examines key musical attributes including tempo, energy, valence, danceability, and popularity. The analysis explores and contextualizes the bands' stylistic changes within broader music industry trends, identifying correlations and contrasts with mainstream music. This comprehensive examination provides insights into how Coldplay and Metallica have evolved over time, their unique contributions to the music landscape, and their alignment with or divergence from prevailing musical trends.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

library(tidyverse)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

library(readr)
library(tidyverse)

coldplay <- read_csv("data/Coldplay.csv")
metallica <- read_csv("data/metallica.csv")
spotify <- read_rds("data/Broader_Spotify_Info.rds")
billboard_100 <- read_rds("data/charts.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

Coldplay and Metallica boast careers spanning over decades, during which they have made significant contributions to the music industry with their unique styles and consistent evolution. This analysis aims to explore various aspects of their musical journeys by examining their discographies and analyzing how their musical attributes have evolved over time. By leveraging comprehensive data from Spotify, this report will delve into key metrics such as tempo, energy, valence, and danceability. Additionally, the analysis will contextualize these bands' stylistic changes within broader music industry trends, using supplementary data sources such as the Billboard Top 100 charts.To provide a detailed comparison and identify patterns, we will ensure that the analysis only considers studio recordings, filtering out live performances and demo versions. This approach will maintain the focus on like-for-like comparisons, allowing us to draw meaningful insights into each band's musical evolution.

# Data  {-}

The dataset provided for the analysis of Coldplay and Metallica's musical progression includes a range of attributes that capture various aspects of their songs. Each track is characterized by its duration, measured in seconds, and the release_date, which indicates the date the album was released. The album_name provides the name of the album to which the track belongs. The explicit attribute indicates whether the song contains explicit lyrics, which can impact its audience and distribution. The popularity score, ranging from 0 to 100, measures the song's popularity on Spotify.Several attributes describe the musical qualities of the tracks. Acousticness measures the confidence that a track is acoustic, with values ranging from 0.0 to 1.0. Danceability assesses how suitable a track is for dancing, based on elements like tempo and beat strength. Energy reflects the intensity and activity level of the song, a higher energy level indicating more dynamic and engaging tracks. Instrumentalness predicts the likelihood of a track containing no vocal content, with higher values indicating more instrumental tracks. Liveness detects the presence of an audience in the recording, with values above 0.8 suggesting a live performance. The dataset also includes loudness, which measures the perceived loudness of the track in LUFS, providing a more accurate reflection of how humans perceive loudness. Speechiness detects the presence of spoken words, with higher values indicating more speech-like content. Tempo measures the speed of the track in beats per minute (BPM), and valence describes the musical positiveness conveyed by the track, with higher values indicating more positive, happy-sounding tracks. Together, these attributes offer a comprehensive view of the musical characteristics and evolution of Coldplay and Metallica over their long careers.

# Comparison

<!-- The following is a code chunk. It must have its own unique name (after the r), or no name. After the comma follows commands for R which are self-explanatory. By default, the code and messages will not be printed in your pdf, just the output: -->

The two box plots below represent the popularity of each album for both Metallica and Coldplay. The popularity scores for Metallica's albums vary widely. Their earlier albums such as "Master of Puppets," "Ride the Lightning," and "Kill 'Em All" show higher popularity scores with less variability. However, newer albums like "Hardwired to Self-Destruct" show lower popularity scores suggest a greater popularity in the 80s which then see's a decline into the 2000's, though some recent works still maintain considerable popularity. On the other hand, Coldplay shows a different trend with a more consistent popularity score across albums. Albums like "A Rush of Blood to the Head" and "Viva La Vida or Death and All His Friends" have higher popularity scores but over all there is less variability, suggesting a steadier fanbase.

```{r Figure1, echo=FALSE, fig.align='center', fig.cap="Coldplay Album Popularity \\label{Figure1}", fig.ext='png', fig.height=4, fig.width=6, warning=FALSE}

# Using functions to filter out all non-studio songs...

coldplay_studio <- studio_filter(coldplay)

# ... and plotting as box and whiskers

c_plot <- box_plotter(coldplay_studio, "popularity", title = "Popularity by album (Coldplay)")
c_plot

```


```{r Figure2, echo=FALSE, fig.align='center', fig.cap="Metallica Album Popularity \\label{Figure2}", fig.ext='png', fig.height=4, fig.width=6, warning=FALSE}

# Using functions to filter out all non-studio songs...

metallica_studio <- studio_filter(metallica)

# ... and plotting as box and whiskers

m_plot <- box_plotter(metallica_studio, "popularity", title = "Popularity by album (Metallica)")
m_plot
```

To contextualize the musical evolution of Coldplay and Metallica within broader music industry trends, I utilized supplementary data sources, including the Billboard Top 100 charts spanning several decades. This analysis involved comparing the attributes of Coldplay and Metallica's tracks—such as tempo, energy, valence, danceability, and popularity with trends observed in mainstream music over the years all of which is represented in the line graphs below. This comparative approach allowed for a deeper understanding of how Coldplay and Metallica's musical progression aligns with or stands apart from the prevailing trends in popular music, shedding light on their unique contributions to the industry and their ability to adapt to or influence broader musical movements.

```{r Figure3, echo=FALSE, fig.align='center', fig.cap="Evolution of Music over time \\label{Figure3}", fig.height=5, fig.width=6, warning=FALSE, fig.ext='png'}

# First wrangling the Spotify and billboard data a little bit and joining them...

billboard_100 <- billboard_100 %>% 
    rename(name = song)

mainstream_music <- inner_join(spotify, billboard_100, by = c("name", "artist"))
mainstream_music <- mainstream_music %>% 
    rename(release_date = date)

# .. then feeding it to my attributes plotting function

attributes <- band_attributes(metallica_studio, coldplay_studio, mainstream_music, band_name1 = "Metallica", band_name2 = "Coldplay", band_name3 = "Billboard100")
print(attributes)
```

Metallica shows fluctuating but overall high energy levels throughout their career. Coldplay started of their careers making more melow music but shows a rising trend in energy over the years, indicating a shift towards more energetic music in their later works.

Valence for both bands shows interesting trends. Metallica generally has a lower valence compared to Coldplay, likely due to the subject matter of some of their songs. Coldplay's valence has varied but shows a slight upward trend, indicating more positive or emotionally uplifting music in recent years.

Metallica maintains a lower danceability score throughout their career but given that they primarily produce thrash metal this is consistent with their genre and style. Coldplay's danceability shows an increasing trend, aindicating a move towards more pop and rhythm-oriented music in recent albums.

Both bands exhibit changes in song duration over the years. Metallica’s song durations have seen more fluctuation but they generally tend to be much longer than that of Coldplay. However, Coldplay's songs have seen an increase in duration over time, possibly refelcting more complex song structures or varied musical explorations.

Compared to the Billboard 100, Metallica's music stands out with higher energy and lower valence, emphasizing their thrash metal identity while Coldplay’s musical progression aligns more closely with mainstream trends, showing increasing energy and danceability, possibly explaining their more consistent levels popularity.

The heat maps below provide further insights into how different each bands musical elements relate to each other within their discography. This analysis contextualizes the distinctive musical characteristics and evolution of Metallica and Coldplay over time, highlighting their genre-specific traits and potential shifts in musical style and approach across their careers.

```{r Figure4, echo=FALSE, fig.align='center', fig.cap="Heatmap of Metallicas musical attributes\\label{Figure4}", fig.height=3, fig.width=5, warning=FALSE, fig.ext='png'}

m_heatmap <- heat_mapper(metallica_studio, title = "Heatmap for Metallica Songs")
m_heatmap

```

For Metallica, we see that loudness and energy are highly positively correlated, as are acousticness and liveness. On the other hand, danceability and acousticness, danceability and loudness, and instrumentalness and valence have notable negative correlations.This suggests that Metallica's loud and energetic music tends to be less danceable and less acoustic.

```{r Figure5, echo=FALSE, fig.align='center', fig.cap="Heatmap of Coldplays musical attributes\\label{Figure5}", fig.height=3, fig.width=5, warning=FALSE, fig.ext='png'}

c_heatmap <- heat_mapper(coldplay_studio, title = "Heatmap for Coldplay Songs")
c_heatmap

```

For Coldplay, energy and loudness, and acousticness and valence all see positive correlations while danceability and acousticness, and liveness and acousticness see negative correlations. This indicates that Coldplay's more energetic and loud songs tend to be less acoustic, and there is a distinct separation between acoustic and danceable tracks.

# Conclusion

In comparing Metallica and Coldplay, their contrasting musical journeys highlight diverse paths to longevity and relevance in the music industry. Metallica has remained steadfast in their heavy metal roots, maintaining a high-energy profile while adapting their sound within the constraints of their genre. Over their career, they've shown fluctuations in popularity, with early albums experiencing peaks. In contrast, Coldplay has demonstrated a more adaptive style, evolving from softer alternative rock to incorporating energetic and danceable elements, aligning with broader pop trends. Their consistent popularity across their discography underscores their ability to evolve while maintaining widespread appeal. Together, these bands exemplify how artistic evolution and genre adaptation can sustain enduring success in contemporary music.
