---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Cultural Dynamics and Persistence in Baby Naming Trends: An Analysis of Influences and Longevity in the United States"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Joshua-Connor Knapp^[__Contributions:__  \\So, how about that airline food?]"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Flex Offenders, Local Garage Gym, South Africa" # First Author's Affiliation
Email1: "joshuaconnok\\@gmail.com" # First Author's Email address


CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
 This report explores the evolving landscape of baby naming trends in the United States, highlighting the influence of cultural factors such as popular media and prominent public figures. By analyzing data spanning from 1910 to 2014, we examine the persistence of popular names and identify gender-specific patterns. Our findings reveal a significant decline in the persistence of boys' names since the 1990s, contrasted with the resilience of girls' names. Additionally, we investigate the impact of actor names on naming conventions using regression models, uncovering substantial correlations between popular actor names and increased name counts in movie titles. These insights underscore the interplay between cultural dynamics and societal preferences in shaping naming trends.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

library(tidyverse)
library(readr)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))



baby_names <- read_rds("data/Baby_Names_By_US_State.rds")
charts <- read_rds("data/charts.rds")
hbo_titles <- read_rds("data/HBO_titles.rds")
hbo_credits <- read_rds("data/HBO_credits.rds")
```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

The realm of baby naming trends in the United States has evolved significantly over the years, influenced by a myriad of factors ranging from popular culture to historical events. Understanding these influences not only provides insights into societal trends but also reflects broader cultural shifts. As we delve into the data, we observe intriguing patterns such as the impact of popular movie characters, prominent political figures, celebrities, and even musical sensations on naming practices. Moreover, analyzing the longevity of these trends reveals whether certain names endure over generations or whether others quickly fade into obscurity. Interestingly, recent findings suggest that while overall naming trends may be stabilizing, gender-specific analyses unveil distinct patterns influenced by factors like actor names. This intersection of data analytics and cultural influence offers a fascinating lens through which to explore the dynamic landscape of baby naming in the US.

# Data  {-}

This report primarily considers, three data sets. All take the form of panel data. The first details baby naming conventions, providing the the count of a particular name for years ranging from 1910 to 2014. The other two data sets are comprised of actor credits and titles released by HBO. Given that both of these data sets shared an ID column referring to a specific title, these data sets were joined into a single. It details, a titles release year, its name, who acted in it, the character they played, its popularity, review score, and so on.  

<!-- The following is a code chunk. It must have its own unique name (after the r), or no name. After the comma follows commands for R which are self-explanatory. By default, the code and messages will not be printed in your pdf, just the output: -->

# Naming Conventions {-}

In examining baby naming trends in the United States, one intriguing aspect is the persistence of popular names over time. The agency has tasked me with analyzing the rank correlation, specifically using Spearman rank correlation, between each year's 25 most popular boys' and girls' names and those of the subsequent three years. This approach allows us to assess whether today's favored names continue to dominate in future years, providing insights into the longevity of naming trends.As such, this investigation aims to confirm or refute the agency's suspicion that since the 1990s, popular name trends have shown reduced persistence compared to earlier decades.

```{r,  warning =  FALSE, fig.align = 'center', fig.cap = "Spearman Correlations Over Time.\\label{Figure1}", fig.ext = 'png', fig.height = 5, fig.width = 8}

# Getting the plot data

plot_data_M <- spearman_corr(baby_names, gender = "M")
plot_data_F <- spearman_corr(baby_names, gender = "F")

# Plotting the data

plots <- ts_corr_plotter(plot_data_M, plot_data_F)
plots
  
```
\hfill

<!-- hfill can be used to create a space, like here between text and table. -->

As can be seen from the tables above, there is a clear divergence in persistence between boys' and girls' naming trends. Boys' names have exhibited a notable decline in staying power since the 90's when compared to earlier decades. In other words, boys have a faster turnover in naming preferences. In contrast, girls' names have shown greater resilience, consistently maintaining or exceeding their popularity levels from the 1990s. While girls' names experienced fluctuations with occasional peaks, they generally reverted to similar levels over time. When examining both genders collectively, the overall trend reveals a decline in the persistence of popular names across the board. This finding underscores shifting societal influences and perhaps evolving cultural dynamics such as interest in cinematic media.

# Cultural Influences {-}

We see further in the figures below, showing that the top five most popular names overall have all past their peaks for both males and females and that naming conventions, while sticky, do not persist indefinitely. While their are likely many cultural and sociological factors that impact on this change, media is likely one of them. 

```{r,  warning =  FALSE, fig.align = 'center', fig.cap = "Top 5 Male Names Over Time.\\label{Figure2}", fig.ext = 'png', fig.height = 5, fig.width = 8}

# Plotting the top 5 persistently popular names per gender throughout the years

top_male_names <- name_plotter(baby_names, gender = "M")
top_male_names

```

```{r,  warning =  FALSE, fig.align = 'center', fig.cap = "Top 5 Female Names Over Time.\\label{Figure3}", fig.ext = 'png', fig.height = 5, fig.width = 8}

# Plotting the top 5 persistently popular names per gender throughout the years

top_female_names <- name_plotter(baby_names, gender = "F")
top_female_names

```
\hfill

<!-- hfill can be used to create a space, like here between text and table. -->

We examine this by first joining the baby names data set to the actors and titles data set according to actor name and the titles release year. This gives us the count of a particular name in accordance with the name of the actor in the title for that titles release year. Using this data set, a dummy variable was created to indicate whether an actor with one of the top 5 most popular names for males and females was present. Subsequently, this dummy variable, along with a titles popularity and review score, was regressed using on the count of a name in the titles release year using Ordinary Least Squares (POLS) to examine its impact. 

```{r, results = 'asis'}

actor_names_M <- names_reg(baby_names, hbo_credits, hbo_titles, gender = "M")
actor_names_M

```

\hfill

<!-- hfill can be used to create a space, like here between text and table. -->

The Breusch-Pagan test results indicate significant heteroscedasticity, suggesting variability in error terms across observations, reinforcing the need for random effects or fixed effects models to account for such variance. Once accounted for, the results for male names highlight significant coefficients across all models, all of which indicate that an actor having one of the top 5 names greatly impacts on name counts for those names. The Hausman test further indicates that the Fixed Effects model is preferable due to its lower P-value, suggesting that unobserved heterogeneity is indeed present and should be accounted for.

```{r, results = 'asis'}

actor_names_F <- names_reg(baby_names, hbo_credits, hbo_titles, gender = "F")
actor_names_F

```

\hfill

<!-- hfill can be used to create a space, like here between text and table. -->

We see similar results for female names, however in this case the Hausman test results provide a larger P-value, indicating that endogeneity may be present and that the Random Effects (RE) for controlling unobserved heterogeneity in this analysis. Despite this, we still see a large positive impact on name counts when actress's have one of the top 5 names. Overall, for both male and female names, these findings underscore the substantial impact of actor name popularity naming conventions, providing valuable insights into consumer preferences.


# Conclusion

In conclusion, our exploration into baby naming trends in the United States has unveiled a dynamic landscape influenced by a multitude of cultural and societal factors. Our analysis has shown that while overall naming trends appear to be stabilizing, there are distinct gender-specific patterns: boys' names exhibit a noticeable decline in persistence since the 1990s, whereas girls' names demonstrate greater resilience over time. These trends reflect evolving cultural dynamics and possibly changing preferences influenced by various media and cultural icons.

Moreover, our investigation into the impact of actor names on baby naming conventions has revealed compelling insights. By linking baby names data with actor credits and title information, we observed significant correlations between the presence of top 5 actor names and increased name counts in movie titles. The adoption of Fixed Effects models, supported by the Hausman test, underscores the importance of accounting for unobserved heterogeneity in such analyses, enhancing the validity of our findings.

Overall, these findings underscore the intricate interplay between cultural influences, media representation, and societal preferences in shaping baby naming trends. They provide valuable insights for understanding consumer behavior and cultural dynamics in contemporary society, highlighting the enduring relevance of popular culture in influencing naming practices.
