---
title: "Country Aid to Ukraine Summary"
author: "Joshus-Connor Knapp"
date: "2024-06-19"
output: pdf_document
---

```{r include=FALSE}

rm(list = ls()) 
gc()

library(tidyverse)
library(readr)

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

alloc <- read_csv("data/Financial Allocations.csv")
commit <- read_csv("data/Financial Commitments.csv")
```


```{r include=FALSE}

# Just a small bit of wrangling...

# ... by joining the data sets by "country" and "EU member" columns so that I only have to work with one data set...

data <- inner_join(alloc, commit, by = c("Country", "EU member"))

# ... and then filtering out countries who are not EU Members

data <- data %>% 
    filter(`EU member` == 1) %>%
    mutate(Allocations_GDP_Percent = (`Total bilateral allocations($ billion)` / `GDP in 2021($ billion)`) * 100) %>% 
    mutate(Commitments_GDP_Percent = (`Total bilateral commitments($ billion)` / `GDP in 2021($ billion)`) * 100)

```

# Summary:

* As can be seen in Figure \ref{Figure2} below, while most EU members are allocating at least as much aid to Ukraine as what they committed, 48% still fail to meet
their commitment. 

* Countries closer to the conflict have the highest allocations, these including mostly Slavic nations and former USSR occupants.

* It is possible that those further from the conflict are naturally more detached from the reality of what is going on in Ukraine. 

## Data Visualisation:

```{r Figure2, echo=FALSE, fig.align='center', fig.cap="Allocation vs. Commitment.\\label{Figure2}", fig.ext='png', fig.height=4, fig.width=6, warning=FALSE}

# Filtering out Estonia and Malta...

data <- data %>%
    filter(!Country %in% c("Estonia", "Malta"))

# ... and plotting the filtered data using my double_bar_plotter function...

compare_plot <- double_bar_plotter(data)
compare_plot

```


```{r Figure3, echo=FALSE, fig.align='center', fig.cap="Proportion who allocate as much as they commit.\\label{Figure3}", fig.ext='png', fig.height=4, fig.width=6, warning=FALSE}

# Seeing how many EU members allocated as much or more than they committed versus how many fell short...

proportions <- allocation_proportions(data)
proportions

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

# Seeing which countries are the relative largest contributers using my "allocation_head" function...

allocation_head <- allocation_head(data)
allocation_head

```
