---
title: "Comparative Analysis of Olympic Success"
author: "Joshua-Connor Knapp"
date: "2024-06-19"
output: pdf_document
---

```{r include=FALSE}

rm(list = ls()) 
gc()

library(tidyverse)
library(readr)

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

Winter <- read_rds("data/winter.rds")
Summer <- read_rds("data/summer.rds")
colnames(Summer)[6] <- "Code"
colnames(Winter)[6] <- "Code"
GDP <- read_rds("data/GDP.rds")

```

# Findings Summary:

 * India has fared somewhat poorly compared to other countries in past Olympic events. This is revealed by the bar graph in Figure \ref{Figure1} below which plots
 the cumulative medals that Brazil, Pakistan, India, and South Africa won during years in which they all competed. These countries were chosen as Pakistan is of
 similar economic standing and Brazil and South Africa are both emerging market economies, one from South America and one from Africa, providing a fair and diverse
 comparison.India falls short of all of them with regard to bronze, silver, and gold medals, except Pakistan. 
 
 * The three most consistently dominant countries, in both the summer and the winter Olympics are France, the United Kingdom, and the United States, all of which
 see medal counts higher than average across the board. This is detailed in Figure \ref{Figure2} below.
 
 * Countries who manage to punch above their weight include Finland, France, Hungary, and Italy, as detailed in the graphs in Figure \ref{Figure3} below. We define
 this as a countries ability to win more medals than the average fairly consistently or at least more often than the other countries. as well as by the number of
 athletes they have competing in a given year. An athlete count below the average futher contributes to this ability.
 
 * My favorite event at the Olympics is athletics. To my disdain, as shown in Figure \ref{Figure4} below, the best athletes in the world come from the United
 States, at least when measuring for the quantity of medal won. They manage to more than double their runner up, the United Kingdom.
 
 \newpage
 
## Data Visualisation:

```{r Figure1, echo=FALSE, fig.align='center', fig.cap="India's Past Olympic Performance. \\label{Figure1}", fig.ext='png', fig.height=3, fig.width=6, message=FALSE, warning=FALSE}

# Joining the summer Olympic and GDP data by country code

summer_gdp <- left_join(Summer, GDP, by = "Code")

# Using my medal counter function

summer_medals <- team_adjuster(summer_gdp)

# Now plotting this data as a country comparison, comparing India to similar and emerging economies...

india_compare <- country_comparison_bar(summer_medals, GDP, "India")
india_compare

```
\hfill

```{r Figure2, echo=FALSE, fig.align='center', fig.cap="Dominant Countries.\\label{Figure2}", fig.ext='png', fig.height=3, fig.width=6, message=FALSE, warning=FALSE}

# A small bit of wrangling, bindin the summer and winter Olympics sets together and join GDP to this new data set by the "Code" column

olympics <- rbind(Summer, Winter)
olympics_gdp <- left_join(olympics, GDP, by = "Code")

# Using the my "team_adjuster" function to account for duplicate medals...

team_adjusted <- team_adjuster(olympics_gdp)

# ... and inseting the output in to my "high_performance" function

dominant_countries <- high_performance(team_adjusted)
dominant_countries

```

```{r Figure3, echo=FALSE, fig.align='center', fig.cap="Countries that punch above their weight.\\label{Figure3}", fig.ext='png', fig.height=4, fig.width=6, message=FALSE, warning=FALSE}

no_odds <- above_weight(summer_gdp, summer_medals)
no_odds

```

```{r Figure4, echo=FALSE, fig.align='center', fig.cap="Countries According to Athletics Medals. \\label{Figure4}", fig.ext='png', fig.height=4, fig.width=6, warning=FALSE}

best_athletes <- all_time_best(olympics_gdp)
best_athletes

```
